---
title: "vignette_presentation"
output: html_document
date: '2022-08-25'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = getwd())
```

<!-- ####################################################################### -->

```{r}
devtools::load_all()
```


## Introduction

### Presentation of the package

The _rtauargus_ package offers an **R** interface for **&tau;-Argus**.
It allows to :

- create inputs (asc, tab and rda files) from R data;
- generate the sequence of instructions to be executed in batch mode (arb file);
- launch a batch &tau;-Argus ;
- retrieve the results in R.

The syntax of some of the arguments closely matches the _batch_ syntax of
&tau;-Argus. This allows a large number of functions to be used without
multiplying the arguments of the functions. The package will also be able to adapt
more easily to possible modifications of the software (new
methods available, additional options...).
The syntax rules for writing batch are given in the &tau;-Argus reference manual and will be specified in a dedicated help section.

> The package was developed on the basis of open source versions of &tau;-Argus (versions 4.2 and above), in particular the latest version available at the time of development (4.2.2b1).
>
> It is not compatible with version 3.5.**_

### Purpose of this document

This document aims to explain the main functionalities of the package,
using relatively simple examples.
A detailed documentation of any function (exhaustive list of arguments,
technical aspects...) is available *via* the dedicated help section.

<p style="text-align: right">
  <a href="#TOC" title="Return to summary">summary &uarr;</a>
</p>


<!-- ####################################################################### -->


## Setup

The following setup should be made before the first use (and no longer afterwards)

### &tau;-Argus

_rtauargus_ fonctions using &tau;-Argus requires that the software can be used from the workstation.
The github repository of &tau;-Argus is here: [https://github.com/sdcTools/tauargus](https://github.com/sdcTools/tauargus). 
The latest releases can be downloaded here: [https://github.com/sdcTools/tauargus/releases](https://github.com/sdcTools/tauargus/releases).

### Dependencies

 _rtauargus_ requires some other R packages.Those are the dependencies to install.
 
  ```{r, echo = FALSE, comment = "•"}
imports <- packageDescription("rtauargus", fields = "Imports")
cat(gsub(", ", "\n", imports))
```
The package _rtauargus_ can be installed now.

<p style="text-align: right">
  <a href="#TOC" title="Return to summary">sommaire &uarr;</a>
</p>


<!-- ####################################################################### -->

## Quick Start

This section explains how to perform a minimal configuration of the package and
how to apply suppressive methods in a single instruction.

### Location of &tau;-Argus

When loading the package, the console displays some information:

```{r library}
library(rtauargus)
```

In particular, a plausible location for the &tau;-Argus software is
predefined. This can be changed for the duration of the R session, as follows:

```{r opt_exe, include = FALSE}
loc_tauargus <- "Y:/Logiciels/TauArgus/TauArgus4.2.2b1/TauArgus.exe"

options(rtauargus.tauargus_exe = loc_tauargus)
```


With this small adjustment done, the package is ready to be used.

> _For a more customized configuration, see the specific vignettes_

### Protecting a single tabular data with `tab_rtauargus()` function

The `tab_rtauargus()` function performs a full processing to protect the table and retrieves the results immediately in R.

Completely abstracting from the inner workings of &tau;-Argus, it allows the entire processing to be made in a single instruction. All intermediate files are created in a local directory.

`tab_rtauargus()` requires the following arguments :

- `tabular`: a data.frame containing the table ;
- `dir_name`: the directory for the outputs;
- `files_name`: all the &tau;-Argus files will be named with it (different extensions);
- `explanatory_vars`: the name of all explanatory variables in `tabular`;
- `secret_var` or `safety_rules`: the way to apply primary suppression (explain later)
- `totcode`: the code for total of each explanatory variable in `tabular`

All the arguments and their default options will be detailed ( where?). 

#### Minimal example

For the following demonstration, a fictitious table will be used:

```{r data}
act_size <-
  data.frame(
    ACTIVITY = c("01","01","01","02","02","02","06","06","06","Total","Total","Total"),
    SIZE = c("tr1","tr2","Total","tr1","tr2","Total","tr1","tr2","Total","tr1","tr2","Total"),
    VAL = c(100,50,150,30,20,50,60,40,100,190,110,300),
    N_OBS = c(10,5,15,2,5,7,8,6,14,20,16,36),
    MAX = c(20,15,20,20,10,20,16,38,38,20,38,38)
  )
act_size
```

As primary rules, we use the two following ones:  

- The n-k dominance rule with n=1 and k = 85
- The minimum frequency rule with n = 3 and a safety range of 10.

To get the results for the dominance rule, we need to specify the largest contributor to each cell, corresponding to the `MAX` variable in the tabular data.  

```{r rtauargus_ex1}
ex1 <- tab_rtauargus(
  act_size,
  dir_name = "tauargus_files",
  files_name = "ex1",
  explanatory_vars = c("ACTIVITY","SIZE"),
  safety_rules = "FREQ(3,10)|NK(1,85)",
  value = "VAL",
  freq = "N_OBS",
  maxscore = "MAX",
  totcode = c(ACTIVITY="Total",SIZE="Total")
)
```

By default, the function displays in the console the logbook content in which 
user can read all steps run by &tau;-Argus. This can be retrieved in the logbook.txt file. With `verbose = FALSE`, the function can be silenced.

By default, the function returns the original dataset with one variable more, 
called `Status`, directly resulting from &tau;-Argus and describing the status of 
each cell as follows:  

-`A`: primary secret cell because of frequency rule;  
-`B`: primary secret cell because of dominance rule (1st contributor);  
-`C`: primary secret cell because of frequency rule (more contributors in case when n>1);  
-`D`: secondary secret cell;  
-`V`: valid cells - no need to mask.

```{r}
ex1
```


All the files generated by the function are written in the specified directory
(`dir_name` argument). The default format for the protected table is csv but it can be changed. All the &tau;-Argus files (.tab, .rda, .arb and .txt) are written in the 
same directory, too. To go further, you can consult the latest version of the &tau;-Argus manual is downloadable here: 
[https://research.cbs.nl/casc/Software/TauManualV4.1.pdf](https://research.cbs.nl/casc/Software/TauManualV4.1.pdf).  


#### Example with hierarchy and primary secret already done

##### Data

For this example, we'd like to protect a table in which companies' turnover 
is broken down by business sectors and size. To load the data, do:

```{r load_data}
data("turnover_act_size")
head(turnover_act_size)
```

The meaning of each variable is:  

-`ACTIVITY`: business sector, hierarchical variables with three levels described
in the `activity_corr_table` dataset. The root is noted "Total";  
-`SIZE`: size of the companies (Number of employees in three categories
+ overall category "Total");  
-`N_OBS`: Frequency, number of companies;  
-`TOT`: turnover value in euros;  
-`MAX`: turnover of the company which contributes the most to the cell.  

##### Hierarchy's file

Before performing the `tab_rtauargus()` function, we have to prepare the hierarchical
information into the appropriate format for &tau;-Argus, *.i.e.* a `.hrc` file.
From a correspondence table, the `write_hrc2()` function does the job for you.

Here, the correspondence table describes the nesting of the three levels of business 
sectors, from the most aggregated to the least one:   

```{r}
data(activity_corr_table)
head(activity_corr_table)
```


```{r hierarchy}
hrc_file_activity <- write_hrc2(
  corr_table = activity_corr_table,
  file_name = "hrc/activity.hrc"
)
```

##### Primary secret 

In this example, we'll apply the primary secret ourselves, *i.e.* not with the 
help of &tau;-Argus. The idea is to use &tau;-Argus with an apriori file.
For that purpose, we create a boolean variable to specify 
which cells don't comply with the primary secret rules. Using the same rules as before, 
we get:  

```{r prim_secret}
turnover_act_size <- turnover_act_size %>%
  mutate(
    is_secret_freq = N_OBS > 0 & N_OBS < 3,
    is_secret_dom = MAX > TOT*0.85,
    is_secret_prim = is_secret_freq | is_secret_dom
  )
```

##### Running &tau;-Argus

Two arguments has to be added to the `tab_rtauargus()` function:  

-`secret_var`, indicating the name of the variable in `tabular` containing the 
primary secret (apriori) information;  
-`hrc`, indicating the name of the hierarchy file to use for `ACTIVITY` variable.  

Since the primary suppression was already specified, no need to use the arguments,
`safety_rules` and `maxscore`. The first one is set by default to "MAN(10)", so 
as to say that a 10% Interval Protection is applied.

By default, `tab_rtauargus()` runs the Modular method to perform the secondary secret. Here, we choose to use the Optimal method by changing the `suppress` argument.

```{r rtauargus_ex2}
ex2 <- tab_rtauargus(
  turnover_act_size,
  dir_name = "tauargus_files/",
  files_name = "ex2",
  explanatory_vars = c("ACTIVITY","SIZE"),
  value = "TOT",
  freq = "N_OBS",
  secret_var = "is_secret_prim",
  hrc = c(ACTIVITY = hrc_file_activity),
  totcode = c(ACTIVITY="Total",SIZE="Total"),
  suppress = "OPT(1,5)",
  verbose=FALSE,
  unif_hrc = FALSE,
  unif_expl = FALSE
)
```

##### Result

```{r}
str(ex2)
```

```{r}
table(ex2$Status)
```

As we can see in the `table()` results, all the primary secret has the status "B" 
in the output produced by &tau;-Argus. To adjust this, we can do:

```{r}
ex2 %>% 
  mutate(
    Status = dplyr::case_when(
      is_secret_freq ~ "A",
      TRUE ~ Status
    )
  ) %>% 
  dplyr::count(Status)
```


### Protecting several tables at the same time, with `tab_muli_manager()` function

The function `tab_multi_manager()` can deal with a whole set of (linked or not) tables.
It is an iterating process, performing secondary suppression with one table at a time and it ensures that the common cells have the same status. When a common cell is concerned by secondary suppression, it reverberates the secret on each table that shares this common cell. The process ends when the secondary secret is consistent in every tables. See more details in vignettes *Manage the protection of linked tables*

#### Data

For this example, two tables will be used : 

```{r}
data("turnover_act_size")
data("turnover_act_cj")
str(turnover_act_cj)
```

The second tabular dataset provides the turnover of companies broken down by 
business sectors (`ACTIVITY`) and type of company (`CJ`). The latter has three 
categories and the overall category is noted "Total".

Since the two tables share one common explanatory variable (`ACTIVITY`), they can't be treated separately without risk of inconsistent protection.

#### Primary secret

The first step consists in indicating whether each cell complies with the primary rules,
or not. A boolean variable is created, equal to TRUE if the cell doesn't comply.

Here, we use the same rules as previously.

```{r}
list_data_2_tabs <- list(
  act_size = turnover_act_size,
  act_cj = turnover_act_cj
) %>%
  purrr::map(
    function(df){
      df %>%
        mutate(
          is_secret_freq = N_OBS > 0 & N_OBS < 3,
          is_secret_dom = MAX > TOT*0.85,
          is_secret_prim = is_secret_freq | is_secret_dom
        )
    }
  )
```


#### Running &tau;-Argus

Now that the primary secret has been specified for both tables, we can run the process.

```{r}
ex3 <- tab_multi_manager(
  list_tables = list_data_2_tabs,
  list_explanatory_vars = list(
    act_size = c("ACTIVITY", "SIZE"),
    act_cj = c("ACTIVITY", "CJ")
  ),
  hrc = c(ACTIVITY = hrc_file_activity),
  dir_name = "ex3",
  value = "TOT",
  freq = "N_OBS",
  secret_var = "is_secret_prim",
  totcode =  "Total"
)
```

By default, the function uses a wrapper of `tab_rtauargus()` function, called `tab_rtauargus2()`, to apply secondary secret with &tau;-Argus. Many default parameters are set. In particular:  
- Interval Protection fixed to 10;
- Suppress method set to "MOD(1,5,1,0,0);
- Output.

When running, the function displays at each iteration which table is treated.

The vignette *Manage the protection of linked tables* provides a full presentation 
of `tab_multi_manager()` function.

